---
title: "Getting High or Die"
author: "Alex Carriero"
date: '2022-10-21'
output: html_document
---

```{r, warning = F, message = F}
library(readr)
library(tidyverse)
library(psych)
library(MASS)
library(pROC)
library(randomForest)
library(xgboost)
library(rpart)
library(rpart.plot)
library(jtools)
library(class)
library(ISLR)
```
--- title page ---
--- add the tabs ---

--- library not require ---


# Data
```{r}
mushrooms <- read.csv("mushrooms.csv")
```

## Attribute Information: (classes: edible=e, poisonous=p)

$cap-shape$: bell=b,conical=c,convex=x,flat=f, knobbed=k,sunken=s
$cap-surface$: fibrous=f,grooves=g,scaly=y,smooth=s
$cap-color$: brown=n,buff=b,cinnamon=c,gray=g,green=r,pink=p,purple=u,red=e,white=w,yellow=y
$bruises$: bruises=t,no=f
$odor$: almond=a,anise=l,creosote=c,fishy=y,foul=f,musty=m,none=n,pungent=p,spicy=s
$gill-attachment$: attached=a,descending=d,free=f,notched=n
$gill-spacing$: close=c,crowded=w,distant=d
$gill-size$: broad=b,narrow=n
$gill-color$: black=k,brown=n,buff=b,chocolate=h,gray=g, green=r,orange=o,pink=p,purple=u,red=e,white=w,yellow=y
$stalk-shape$: enlarging=e,tapering=t
$stalk-root$: bulbous=b,club=c,cup=u,equal=e,rhizomorphs=z,rooted=r,missing=?
$stalk-surface-above-ring$: fibrous=f,scaly=y,silky=k,smooth=s
$stalk-surface-below-ring$: fibrous=f,scaly=y,silky=k,smooth=s
$stalk-color-above-ring$: brown=n,buff=b,cinnamon=c,gray=g,orange=o,pink=p,red=e,white=w,yellow=y
$stalk-color-below-ring$: brown=n,buff=b,cinnamon=c,gray=g,orange=o,pink=p,red=e,white=w,yellow=y
$veil-type$: partial=p,universal=u
$veil-color$: brown=n,orange=o,white=w,yellow=y
$ring-number$: none=n,one=o,two=t
$ring-type$: cobwebby=c,evanescent=e,flaring=f,large=l,none=n,pendant=p,sheathing=s,zone=z
$spore-print-color$: black=k,brown=n,buff=b,chocolate=h,green=r,orange=o,purple=u,white=w,yellow=y
$population$: abundant=a,clustered=c,numerous=n,scattered=s,several=v,solitary=y
$habitat$: grasses=g,leaves=l,meadows=m,paths=p,urban=u,waste=w,woods=d

```{r}
mushrooms <- read_csv("mushrooms.csv")
mushrooms
```

```{r}
library(tidyverse)
```


```{r}
# a bad poem:  
# there once was a mushroom that we wanted to eat 
# would it be bitter or would it be sweet 

# a machine learning classifier will help us decide 
# if we will get high or if we will die

# soon we will be in the clouds
# or have nasty stuff in our mouths

# liz truss was defeated by the salad

# a leave node shows if its red or blue

# the red ones look nice 
# the blue ones look bad
# if but if i pick red and its poisonous
# i will be sad
```

--- detail about relevant columns ---
--- overkill on tables ---

--- don't needlessly skip simplicity ---

# Simple model

```{r}
print("Hello world")
```

# Improved model

```{r}
# here we are supposed to make a cross validation and maybe some crazy threshold stuff
```

# Something

```{r}
# Judith

# explain method
# assess predictions
# interpret parameters
# if no parameters: interpret contribution of features
# make conclusions about predictions
# use plots
```


# Florian
```{r}
# Load data
mushrooms <- read_csv("~/Documents/GitHub/Getting-High-or-Die/mushrooms.csv")

# check the data
View(mushrooms)
str(mushrooms)

# Tidy the data
mushrooms_tidy <- mushrooms %>%
  mutate(across(where(is.character),as_factor)) %>% 
  select(-`veil-type`) # there is only one cat 

# change the - to _ else the random forest will trip
names(mushrooms_tidy) <- gsub(x = names(mushrooms_tidy), pattern = "-", replacement = "_") 

# look at the data
describe(mushrooms_tidy)

# SPLIT THE DATA

#make this example reproducible
set.seed(0)

mushrooms_tidy_split <- 
  mushrooms_tidy %>% 
  mutate(split = sample(rep(c("train", "test"), times = c(6500, 1624))))

mushrooms_train <- 
  mushrooms_tidy_split %>% 
  filter(split == "train") %>%
  dplyr::select(-split)

mushrooms_test <- 
  mushrooms_tidy_split %>% 
  filter(split == "test") %>% 
  dplyr::select(-split)

# Simple model 2 logistig regression
M1a <- glm(class ~ ., data = mushrooms_train, family = "binomial")
summary(M1a)

pred_prob <- predict(M1a, newdata = mushrooms_test,type = "response")
pred_lr   <- factor(pred_prob > .5)
table(true = mushrooms_test$class, predicted = pred_lr)

roc_lr1 <- roc(mushrooms_test$class, pred_prob)
ggroc(roc_lr1) + theme_minimal() + labs(title = "LR1")

# Simple model 2 LDA
M2a <- lda(class ~ ., data = mushrooms_train)
pred_lda <- predict(M2a, newdata = mushrooms_test)
table(true = mushrooms_test$class, predicted = pred_lda$class)

# Improved model 1 Tree
M3a <- rpart(class ~ ., data = mushrooms_tidy)
rpart.plot(M3a)

# Improved model 2 Free tree
M3b <- rpart(class ~ ., data = mushrooms_tidy, control = rpart.control(minbucket = 1, cp = 0)) 
rpart.plot(M3b)

# Improved model 3 Random forest
M3c <- randomForest(class ~ ., data = mushrooms_tidy)
M3c

# Improved model 4 XGboost

# select the predictors and outcomes
mushrooms_train_x <- mushrooms_train %>% 
  dplyr::select(-class)
mushrooms_train_y <- mushrooms_train %>% 
  dplyr::select(class)
  
mushrooms_test_x <- mushrooms_test %>% 
  dplyr::select(-class)
mushrooms_test_y <- mushrooms_test %>% 
  dplyr::select(class)

#define final training and testing sets
xgb_train = xgb.DMatrix(data = data.matrix(mushrooms_train_x), label = data.matrix(mushrooms_train_y))
xgb_test = xgb.DMatrix(data =  data.matrix(mushrooms_test_x), label =  data.matrix(mushrooms_test_y))

#define watchlist
watchlist = list(train=xgb_train, test=xgb_test)

#fit XGBoost model and display training and testing data at each round
model = xgb.train(data = xgb_train, max.depth = 3, watchlist=watchlist, nrounds = 100)

# the RSME starts to increase after 96
# define final model
final = xgboost(data = xgb_train, max.depth = 3, nrounds = 96, verbose = 0)

```

# Christoph

```{r}
<<<<<<< HEAD
mushrooms <- read_csv("mushrooms.csv")
mushrooms
=======
mushrooms <- read.csv("mushrooms.csv")
>>>>>>> bab9fe1f78ab03fa892238ab13a50ea0e4ca63c5
```

```{r}
# data manipulation
names(mushrooms) <- gsub(x = names(mushrooms), pattern = "\\-", replacement = "_")  
mushrooms <- na.exclude(mushrooms)
mushrooms <- mushrooms %>% mutate_if(is.character,as.factor)

split <- c(rep("train", round(nrow(mushrooms)*0.7,0)), rep("valid", round(nrow(mushrooms)*0.3,0)))
mushrooms <- mushrooms %>% mutate(split = sample(split))

mushrooms_train <- mushrooms %>% filter(split == "train") %>% dplyr::select(-split,-veil_type)
mushrooms_valid <- mushrooms %>% filter(split == "valid") %>% dplyr::select(-split,-veil_type)
```

```{r}
# all predictors. But apparently model doesn't converge
fit0 <- glm(class ~ ., family = binomial, data=mushrooms_train)

# Train data table
prob_lr <- predict(fit0, type = "response")
pred_lr <- ifelse(prob_lr > .5, 1, 0)

table(true = mushrooms_train$class, pred = pred_lr)

# Validation data table
pred_prob <- predict(fit0, newdata = mushrooms_valid, type = "response")
pred_lr   <- factor(pred_prob > .5, labels = c("No", "Yes"))

table(true = mushrooms_valid$class, predicted = pred_lr)
```

```{r}
# model with many predictors, mainly used in my analysis
fit1 <- glm(class ~ cap_shape+cap_surface+habitat+ring_type+gill_spacing+gill_size+stalk_shape+stalk_surface_above_ring+stalk_surface_below_ring, family = binomial, data=mushrooms_train)
#summary(fit1)
#summ(fit1, exp=T)

# Train data table
prob_lr <- predict(fit1, type = "response")
pred_lr <- ifelse(prob_lr > .5, 1, 0)

table(true = mushrooms_train$class, pred = pred_lr)

# Validation data table
pred_prob <- predict(fit1, newdata = mushrooms_valid, type = "response")
pred_lr   <- factor(pred_prob > .5, labels = c("No", "Yes"))

table(true = mushrooms_valid$class, predicted = pred_lr)
```

### Example with Fit1 only

```{r}
# visualisation of probability for edible/poisonous
tibble(observed  = mushrooms_train$class, 
       predicted = predict(fit1, type = "response")) %>% 
  ggplot(aes(y = predicted, x = observed, colour = observed)) +
  geom_point(position = position_jitter(width = 0.2), alpha = .3) +
  scale_colour_manual(values = c("dark blue", "orange"), guide = "none") +
  theme_minimal() +
  labs(y = "Predicted probability to be poisonous")
```

```{r}
# for loop to test different threshold values
thres <- data.frame(seq(0.05,0.95,length.out=19))
TPR <- data.frame(matrix(nrow = length(thres), ncol = 1))
TNR <- data.frame(matrix(nrow = length(thres), ncol = 1))
ACC <- data.frame(matrix(nrow = length(thres), ncol = 1))
FPR <- data.frame(matrix(nrow = length(thres), ncol = 1))
PPV <- data.frame(matrix(nrow = length(thres), ncol = 1))
NPV <- data.frame(matrix(nrow = length(thres), ncol = 1))

for(i in 1:nrow(thres)){
  pred_prob <- predict(fit1, type = "response")
  pred_lr   <- factor(pred_prob > thres[i,1], labels = c("No", "Yes"))
  cmat_lr <- table(true = mushrooms_train$class, predicted = pred_lr)
  TN <- cmat_lr[1, 1]
  FN <- cmat_lr[2, 1]
  FP <- cmat_lr[1, 2]
  TP <- cmat_lr[2, 2]
  ACC[i,1] <- (TP + TN) / sum(cmat_lr)
  TPR[i,1] <- TP / (TP + FN)
  TNR[i,1] <- TN / (TN + FP)
  FPR[i,1] <- FP / (TN + FP)
  PPV[i,1] <- TP / (TP + FP)
  NPV[i,1] <- TN / (TN + FN)
}
```

```{r}
# data manipulation to have all estimates in one dataframe
alpha <- seq(0.05,0.95,length.out=19)

TPR$alpha <- alpha
TNR$alpha <- alpha
ACC$alpha <- alpha
FPR$alpha <- alpha
PPV$alpha <- alpha
NPV$alpha <- alpha

TPR$TPR <- rep("TPR", nrow(TPR))
TNR$TNR <- rep("TNR", nrow(TNR))
ACC$ACC <- rep("ACC", nrow(ACC))
FPR$FPR <- rep("FPR", nrow(FPR))
PPV$PPV <- rep("PPV", nrow(PPV))
NPV$NPV <- rep("NPV", nrow(NPV))

colnames(TPR) <- c("Prob", "alpha","Facet")
colnames(TNR) <- c("Prob", "alpha","Facet")
colnames(ACC) <- c("Prob", "alpha","Facet")
colnames(FPR) <- c("Prob", "alpha","Facet")
colnames(PPV) <- c("Prob", "alpha","Facet")
colnames(NPV) <- c("Prob", "alpha","Facet")

comb <- rbind(TPR,TNR,ACC,FPR,PPV,NPV)
con_res <- cbind(alpha, ACC[1],TPR[1],TNR[1],FPR[1],PPV[1],NPV[1])
colnames(con_res) <- c("alpha", "ACC","TPR","TNR", "FPR","PPV","NPV")

mean_prop <- comb %>% filter(Facet == "TPR" | Facet == "TNR") %>% group_by(alpha) %>%
  summarise(mean(Prob)) 
mean_prop # .3 is the best probability
```

```{r}
# plotting the different estimates
comb %>% 
  filter(!Facet == "FPR") %>%
    ggplot(aes(x= alpha, y= Prob ,fill = Facet, color=Facet)) + 
    geom_line() +
    theme_minimal() +
    theme(plot.title = element_text(size = rel(2), face = "bold")) +
    labs(title = "Figure 1. Threshold to choose", 
       x = "Threshold", 
       y = "Proportion of correct") 

# conf_matrix with the newly chosen threshold
pred_prob <- predict(fit1, type = "response")
pred_lr   <- factor(pred_prob > .3, labels = c("No", "Yes"))

table(true = mushrooms_train$class, predicted = pred_lr)
```

### Validation data example
```{r}
# Validation data table
pred_prob <- predict(fit1, newdata = mushrooms_valid, type = "response")
pred_lr   <- factor(pred_prob > .5, labels = c("No", "Yes"))

table(true = mushrooms_valid$class, predicted = pred_lr)
```

```{r}
TPR <- data.frame(matrix(nrow = length(thres), ncol = 1))
TNR <- data.frame(matrix(nrow = length(thres), ncol = 1))
ACC <- data.frame(matrix(nrow = length(thres), ncol = 1))
FPR <- data.frame(matrix(nrow = length(thres), ncol = 1))
PPV <- data.frame(matrix(nrow = length(thres), ncol = 1))
NPV <- data.frame(matrix(nrow = length(thres), ncol = 1))

for(i in 1:nrow(thres)){
  pred_prob <- predict(fit1, newdata = mushrooms_valid, type = "response")
  pred_lr   <- factor(pred_prob > thres[i,1], labels = c("No", "Yes"))
  cmat_lr <- table(true = mushrooms_valid$class, predicted = pred_lr)
  TN <- cmat_lr[1, 1]
  FN <- cmat_lr[2, 1]
  FP <- cmat_lr[1, 2]
  TP <- cmat_lr[2, 2]
  ACC[i,1] <- (TP + TN) / sum(cmat_lr)
  TPR[i,1] <- TP / (TP + FN)
  TNR[i,1] <- TN / (TN + FP)
  FPR[i,1] <- FP / (TN + FP)
  PPV[i,1] <- TP / (TP + FP)
  NPV[i,1] <- TN / (TN + FN)
}
```

```{r}
alpha <- seq(0.05,0.95,length.out=19)

TPR$alpha <- alpha
TNR$alpha <- alpha
ACC$alpha <- alpha
FPR$alpha <- alpha
PPV$alpha <- alpha
NPV$alpha <- alpha

TPR$TPR <- rep("TPR", nrow(TPR))
TNR$TNR <- rep("TNR", nrow(TNR))
ACC$ACC <- rep("ACC", nrow(ACC))
FPR$FPR <- rep("FPR", nrow(FPR))
PPV$PPV <- rep("PPV", nrow(PPV))
NPV$NPV <- rep("NPV", nrow(NPV))

colnames(TPR) <- c("Prob", "alpha","Facet")
colnames(TNR) <- c("Prob", "alpha","Facet")
colnames(ACC) <- c("Prob", "alpha","Facet")
colnames(FPR) <- c("Prob", "alpha","Facet")
colnames(PPV) <- c("Prob", "alpha","Facet")
colnames(NPV) <- c("Prob", "alpha","Facet")

comb_val <- rbind(TPR,TNR,ACC,FPR,PPV,NPV)
con_res_val <- cbind(alpha, ACC[1],TPR[1],TNR[1],FPR[1],PPV[1],NPV[1])
colnames(con_res_val) <- c("alpha", "ACC","TPR","TNR", "FPR","PPV","NPV")


mean_prop <- comb %>% filter(Facet == "TPR" | Facet == "TNR") %>% group_by(alpha) %>%
  summarise(mean(Prob)) 
mean_prop # .3 is the best probability
```

```{r}
comb %>% 
  filter(!Facet == "FPR") %>%
    ggplot(aes(x= alpha, y= Prob ,fill = Facet, color=Facet)) + 
    geom_line() +
    theme_minimal() +
    theme(plot.title = element_text(size = rel(2), face = "bold")) +
    labs(title = "Figure 1. Threshold to choose", 
       x = "Threshold", 
       y = "Proportion of correct") 

pred_prob <- predict(fit1, newdata = mushrooms_valid, type = "response")
pred_lr   <- factor(pred_prob > .3, labels = c("No", "Yes"))

table(true = mushrooms_valid$class, predicted = pred_lr)
```
### LDA
```{r}
lda_mod <- lda(class ~ cap_shape+cap_surface+habitat+ring_type+gill_spacing+gill_size+stalk_shape+stalk_surface_above_ring+stalk_surface_below_ring, data = mushrooms_train)

lda_mod
```

```{r}
# train data
pred_lda <- predict(lda_mod)
cmat_lda <- table(true = mushrooms_train$class, pred = pred_lda$class)
PPV <- cmat_lda[2, 2] / sum(cmat_lda[, 2])
NPV <- cmat_lda[1, 1] / sum(cmat_lda[, 1])
lda_t <- cbind(PPV,NPV)

pred_lda_val <- predict(lda_mod, newdata = mushrooms_valid)
cmat_lda_new <- table(true = mushrooms_valid$class, predicted = pred_lda_val$class)
PPV_v <- cmat_lda_new[2, 2] / sum(cmat_lda_new[, 2])
NPV_v <- cmat_lda_new[1, 1] / sum(cmat_lda_new[, 1])

lda_v <- cbind(PPV_v,NPV_v)
lda <- cbind(lda_t,lda_v)
```

```{r}
con_res_lr <- con_res %>% filter(alpha == 0.3 | alpha == 0.6) %>% dplyr::select(alpha, PPV, NPV)
con_res_val_lr <- con_res_val %>% filter(alpha == 0.3 | alpha == 0.6) %>% dplyr::select(PPV, NPV)
colnames(con_res_val_lr) <- c("PPV_v","NPV_v")
lr <- cbind(con_res_lr,con_res_val_lr)
lda
lr
```

### Brier Score
```{r}
pred_prob <- predict(fit1, newdata = mushrooms_valid, type = "response")
mean((pred_prob - (as.numeric(mushrooms_valid$class) - 1)) ^ 2)
# the mean squared difference between the probability and the true class is 0.0277
```
### ROC curve

```{r}
prob_lr1 <- predict(fit1, type = "response")
prob_lr2 <- predict(fit2, type = "response")
```

```{r}
roc_lr1 <- roc(mushrooms_train$class, prob_lr1)
roc_lr2 <- roc(mushrooms_train$class, prob_lr2)
ggroc(roc_lr1) + theme_minimal() + labs(title = "LR1")
ggroc(roc_lr2) + theme_minimal() + labs(title = "LR2")
roc_lr1
roc_lr2
```

# Al


```{r}
data<- read.csv("mushrooms.csv")
```

```{r}
# tidy data -- merge redundant categories 
mushrooms <- data %>% 
    mutate(across(where(is.character), as_factor))%>%
    dplyr::select(-veil.type, -stalk.root)%>%     
    mutate(cap.shape = cap.shape %>% 
             fct_recode("bell"    = "b",
                        "conical" = "c",
                        "convex"  = "x",
                        "flat"    = "f", 
                        "knobbed" = "k",
                        "sunken"  = "s"),
           cap.surface = cap.surface %>% 
             fct_recode("fibrous" = "f",
                        "grooves" = "g",
                        "scaly"   = "y",
                        "smooth"  = "s"), 
           bruises = bruises %>% 
             fct_recode("true"       = "t",
                        "false"      = "f"), 
           odor = odor %>%
             fct_recode("almond"     = "a",
                        "anise"      = "l",
                        "creosote"   = "c",
                        "fishy"      = "y",
                        "foul"       = "f",
                        "musty"      = "m",
                        "none"       = "n",
                        "pungent"    = "p",
                        "spicy"      = "s"),
           cap.color = cap.color %>%                 
             fct_recode("brown"      = "n",
                        "brown"      = "b",
                        "brown"      = "c", 
                        "pink"       = "u",                      
                        "pink"       = "e", 
                        "pink"       = "p", 
                        "gray"       = "g",
                        "green"      = "r", 
                        "white"      = "w",
                        "yellow"     = "y"),
           gill.attachment = gill.attachment %>% 
             fct_recode("attached"   = "a",
                        "free"       = "f"), 
           gill.spacing = gill.spacing %>% 
             fct_recode("close"      = "c",
                        "crowded"    = "w"),
           gill.size = gill.size %>% 
             fct_recode("broad"      = "b",
                        "narrow"     = "n"), 
           gill.color = gill.color %>% 
             fct_recode("black"      ="k",
                        "brown"      ="n",
                        "brown"      ="b",
                        "brown"      ="h",
                        "gray"       ="g", 
                        "green"      ="r",
                        "orange"     ="o",
                        "pink"       ="p",
                        "pink"       ="u",
                        "pink"       ="e",
                        "white"      ="w",
                        "yellow"     ="y"),
           stalk.shape = stalk.shape %>% 
             fct_recode("enlarging"  ="e",
                        "tapering"   ="t"), 
           stalk.surface.above.ring = stalk.surface.above.ring %>% 
             fct_recode("fibrous"    ="f",
                        "scaly"      ="y",
                        "silky"      ="k",
                        "smooth"     ="s"), 
          stalk.surface.below.ring = stalk.surface.below.ring %>% 
             fct_recode("fibrous"    ="f",
                        "scaly"      ="y",
                        "silky"      ="k",
                        "smooth"     ="s"), 
          stalk.color.above.ring = stalk.color.above.ring %>% 
            fct_recode("brown"       ="n",
                       "brown"       ="b",
                       "brown"       ="c",
                       "gray"        ="g",
                       "orange"      ="o",
                       "pink"        ="p",
                       "pink"        ="e",
                       "white"       ="w",
                       "yellow"      ="y",), 
          stalk.color.below.ring = stalk.color.below.ring %>% 
            fct_recode("brown"       ="n",
                       "brown"       ="b",
                       "brown"       ="c",
                       "gray"        ="g",
                       "orange"      ="o",
                       "pink"        ="p",
                       "pink"        ="e",
                       "white"       ="w",
                       "yellow"      ="y",), 
          veil.color = veil.color %>%
            fct_recode("brown"       ="n",
                       "orange"      ="o",
                       "white"       ="w",
                       "yellow"      ="y"),
          ring.number = ring.number %>% 
            fct_recode("none"        ="n",
                       "one"         ="o",
                       "two"         ="t"), 
          ring.type = ring.type %>% 
            fct_recode("evanescent"  ="e",
                       "flaring"     ="f",
                       "large"       ="l",
                       "none"        ="n",
                       "pendant"     ="p"),
          spore.print.color = spore.print.color %>% 
            fct_recode("black"   ="k",
                       "brown"   ="n",
                       "brown"   ="b",
                       "brown"   ="h",
                       "green"   ="r",
                       "orange"  ="o",
                       "purple"  ="u",
                       "white"   ="w",
                       "yellow"  ="y"), 
         population = population %>% 
           fct_recode("group"    ="a",
                      "group"    ="c",
                      "group"    ="n",
                      "group"    ="s",
                      "group"    ="v",
                      "solitary" ="y"), 
         habitat = habitat %>% 
           fct_recode("grasses"  ="g",
                      "woods"    ="l",
                      "grasses"  ="m",
                      "urban"    ="p",
                      "urban"    ="u",
                      "waste"    ="w",
                      "woods"    ="d"))
```

```{r}
# test and train
n = dim(mushrooms)[1]
mushrooms_split <- mushrooms %>% 
                   mutate(split = sample(rep(c("train", "test"), times = c(round(.8*n), round(.2*n)))))

mushrooms_train <- mushrooms_split %>% 
                  filter(split == "train") %>%
                  dplyr::select(-split)

mushrooms_test  <- mushrooms_split %>% 
                  filter(split == "test") %>%
                  dplyr::select(-split)
```

```{r}
# beginner 
beginner       <- mushrooms %>% 
                  dplyr::select(class, cap.shape, cap.color, bruises, gill.color, stalk.shape, ring.number, population, habitat)

beginner_train <- mushrooms_train %>% 
                  dplyr::select(class, cap.shape, cap.color, bruises, gill.color, stalk.shape, ring.number, population, habitat)

beginner_test  <- mushrooms_test %>% 
                  dplyr::select(class, cap.shape, cap.color, bruises, gill.color, stalk.shape, ring.number, population, habitat)
                

# advanced 
advanced       <- mushrooms 
advanced_train <- mushrooms_train
advanced_test  <- mushrooms_test
```

```{r}
lr_b <- glm(class~. , family = "binomial",  data = beginner) # all good
```

```{r}
lr_a <- glm(class~. , family = "binomial",  data = advanced) # no convergence
```

## Christoph & Florian

```{r}
generate_formulas <- function(p, x_vars, y_var) {
  x_vars <- colnames(x_vars)
  # Input checking
  if (p %% 1 != 0)           stop("Input an integer n")
  if (p > length(x_vars))    stop("p should be smaller than number of vars")
  if (!is.character(x_vars)) stop("x_vars should be a character vector")
  if (!is.character(y_var))  stop("y_vars should be character type")
  
  # combn generates all combinations, apply turns them into formula strings
  apply(combn(x_vars, p), 2, function(vars) {
    paste0(y_var, " ~ ", paste(vars, collapse = " + "))
  })
}
```

```{r}
mushrooms_begin_x <- beginner %>%
  dplyr::select(-class)
```

```{r}
formulas_2 <- generate_formulas(p=2,x_vars=mushrooms_begin_x, y_var="class")
formulas_3 <- generate_formulas(p=3,x_vars=mushrooms_begin_x, y_var="class")
formulas_4 <- generate_formulas(p=4,x_vars=mushrooms_begin_x, y_var="class")
formulas_5 <- generate_formulas(p=5,x_vars=mushrooms_begin_x, y_var="class")
formulas_6 <- generate_formulas(p=6,x_vars=mushrooms_begin_x, y_var="class")
```

```{r}
best_model <- function(formulas,train,valid, valid_y){
  out <-data.frame(matrix(nrow = length(formulas), ncol = 2))
  
  for(i in 1:length(formulas)){
    model <- glm(formulas[i], family=binomial, data=train)
    pred_prob <- predict(model, type = "response", newdata = valid)
    roc_lr1 <- roc(valid_y, pred_prob)
    out[i,1] <- roc_lr1$auc[1]
    out[i,2] <- formulas[i]
    }
  colnames(out) <- c("auc", "formula")
  out <- out[order(out$auc,decreasing = T),]
  
return(best=out[1:3, ])
}
```

```{r, echo=FALSE}
pred2 <- best_model(formulas=formulas_2,train=beginner_train,valid=beginner_test,valid_y=beginner_test$class)
pred3 <- best_model(formulas=formulas_3,train=beginner_train,valid=beginner_test,valid_y=beginner_test$class)
pred4 <- best_model(formulas=formulas_4,train=beginner_train,valid=beginner_test,valid_y=beginner_test$class)
pred5 <- best_model(formulas=formulas_5,train=beginner_train,valid=beginner_test,valid_y=beginner_test$class)
pred6 <- best_model(formulas=formulas_6,train=beginner_train,valid=beginner_test,valid_y=beginner_test$class)
```

```{r}
max(pred2$auc)
max(pred3$auc)
max(pred4$auc)
max(pred5$auc)
max(pred6$auc)
# reasonable to take 3-4 predictors
```
```{r}
pred3$formula
pred4$formula
```

```{r}
fit1 <- glm(class ~ bruises + gill.color + habitat, family=binomial, data=beginner_train)
pred_prob <- predict(fit1, type = "response", newdata = beginner_test)
roc <- roc(beginner_test$class, pred_prob)
ggroc(roc) + theme_minimal() + labs(title = pred3$formula[1])
```
```{r}
# Validation data table
pred_lr   <- factor(pred_prob_1 > .5, labels = c("No", "Yes"))
table(true = beginner_test$class, predicted = pred_lr)
```

```{r}
thres <- data.frame(seq(0.05,0.95,length.out=19))
TPR <- data.frame(matrix(nrow = length(thres), ncol = 1))
TNR <- data.frame(matrix(nrow = length(thres), ncol = 1))
ACC <- data.frame(matrix(nrow = length(thres), ncol = 1))
FPR <- data.frame(matrix(nrow = length(thres), ncol = 1))
PPV <- data.frame(matrix(nrow = length(thres), ncol = 1))
NPV <- data.frame(matrix(nrow = length(thres), ncol = 1))

for(i in 1:nrow(thres)){
  pred_prob <- predict(fit1, newdata = beginner_test, type = "response")
  pred_lr   <- factor(pred_prob > thres[i,1], labels = c("No", "Yes"))
  cmat_lr <- table(true = beginner_test$class, predicted = pred_lr)
  TN <- cmat_lr[1, 1]
  FN <- cmat_lr[2, 1]
  FP <- cmat_lr[1, 2]
  TP <- cmat_lr[2, 2]
  ACC[i,1] <- (TP + TN) / sum(cmat_lr)
  TPR[i,1] <- TP / (TP + FN)
  TNR[i,1] <- TN / (TN + FP)
  FPR[i,1] <- FP / (TN + FP)
  PPV[i,1] <- TP / (TP + FP)
  NPV[i,1] <- TN / (TN + FN)
}
```

```{r}
alpha <- seq(0.05,0.95,length.out=19)

TPR$alpha <- alpha
TNR$alpha <- alpha
ACC$alpha <- alpha
FPR$alpha <- alpha
PPV$alpha <- alpha
NPV$alpha <- alpha

TPR$TPR <- rep("TPR", nrow(TPR))
TNR$TNR <- rep("TNR", nrow(TNR))
ACC$ACC <- rep("ACC", nrow(ACC))
FPR$FPR <- rep("FPR", nrow(FPR))
PPV$PPV <- rep("PPV", nrow(PPV))
NPV$NPV <- rep("NPV", nrow(NPV))

colnames(TPR) <- c("Prob", "alpha","Facet")
colnames(TNR) <- c("Prob", "alpha","Facet")
colnames(ACC) <- c("Prob", "alpha","Facet")
colnames(FPR) <- c("Prob", "alpha","Facet")
colnames(PPV) <- c("Prob", "alpha","Facet")
colnames(NPV) <- c("Prob", "alpha","Facet")

comb_val <- rbind(TPR,TNR,ACC,FPR,PPV,NPV)
con_res_val <- cbind(alpha, ACC[1],TPR[1],TNR[1],FPR[1],PPV[1],NPV[1])
colnames(con_res_val) <- c("alpha", "ACC","TPR","TNR", "FPR","PPV","NPV")
```

```{r}
comb_val %>% 
  filter(!Facet == "FPR") %>%
    ggplot(aes(x= alpha, y= Prob ,fill = Facet, color=Facet)) + 
    geom_line() +
    theme_minimal() +
    theme(plot.title = element_text(size = rel(2), face = "bold")) +
    labs(title = "Figure 1. Threshold to choose", 
       x = "Threshold", 
       y = "Proportion of correct") 
```

```{r}
mean_prop <- comb_val %>% filter(Facet == "TPR" | Facet == "TNR" | Facet == "Acc") %>% group_by(alpha) %>%
  summarise(mean(Prob)) 
mean_prop <- mean_prop[order(mean_prop$`mean(Prob)`, decreasing=T),]
mean_prop
```


```{r}
pred_prob <- predict(fit1, newdata = beginner_test, type = "response")
pred_lr   <- factor(pred_prob > .4, labels = c("No", "Yes"))

table(true = beginner_test$class, predicted = pred_lr)
```
